---
title: "Covid91 vaccine study Final2024"
author: "Amia Little"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the
four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$:Covid921 vaccine is not effective in males.

$H_A$:Covid921 vaccine is effective in males.



### Methods

Treatment is CAT
Sex is CAT
So i use the CAT~CAT helper file


### Descriptive Results

#### Graphical Results 

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected+ treatment,data=males, type="percent")

```


#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```




## Females 

$H_0$:Covid921 vaccine is not effective in females.

$H_A$:Covid921 vaccine is effective in females.


### Methods

Treatment is CAT 
Sex is CAT 
So i use the CAT~CAT helper file

### Descriptive Results


#### Graphical Results 

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected+ treatment,data=females, type="percent")

```


#### Numerical Results 

```{r}
table1 <- xtabs(~treatment + infected, data=females)
rowPerc(table1)
colPerc(table1)
```




#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```




## LGBTQ 

$H_0$:Covid921 vaccine is not effective in LGBTQ.

$H_A$:Covid921 vaccine is effective in LGBTQ.

### Methods

### Descriptive Results

#### Graphical Results

#### Numerical Results 

#### Inferential Results




## Druggies 

$H_0$:Covid921 vaccine is not effective in druggies.

$H_A$:Covid921 vaccine is effective in druggies.

### Methods
 
The result is a categorical variable
 
 
### Descriptive Results

#### Graphical Results

#### Numerical Results 

#### Inferential Results



```{r}
males=filter(FinalData,finaldata$sex=="M") 
barchartGC(treatment + infected,data=males) 
barchartGC((~treatment + infected,data=males,data= type="percent")
```




# Overall Results and Conclusions