---
title: "Covid91 vaccine study Final2024"
author: "Amia Little"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the
four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$:Covid921 vaccine is not effective in males.

$H_A$:Covid921 vaccine is effective in males.



### Methods

Treatment is CAT
Sex is CAT
So i use the CAT~CAT helper file


### Descriptive Results

#### Graphical Results 

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected+ treatment,data=males, type="percent")

```

This table shows that males who got the vaccine are less likely to get covid than people who got placebo.

#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected, data=males)
rowPerc(table1)
colPerc(table1)
```

This shows that 74% of males who took the placebo got covid 19 and 26% who took the vaccine got covid 19 as well.

#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Both the chi and fisher test have the same p-value of p-value= 0.000000000000002 which is lower than 0.5 therefore we reject the null hypothesis.


## Females 

$H_0$:Covid921 vaccine is not effective in females.

$H_A$:Covid921 vaccine is effective in females.


### Methods

Treatment is CAT 
Sex is CAT 
So i use the CAT~CAT helper file

### Descriptive Results


#### Graphical Results 

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected+ treatment,data=females, type="percent")

```

This shows that females who took the vaccine had a lower rate of getting covid 19 as opposed to those who took the placebo


#### Numerical Results 

```{r}
table1 <- xtabs(~treatment + infected, data=females)
rowPerc(table1)
colPerc(table1)
```

This shows that 65% of females who took the placebo got covid 19 while 36% of females who got the vaccine comtracted covid 19.



#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The results of the p-value for chi and fishers test is p-value= 0.000000001247 which is lower than 0.5 therefore we reject the null hypothesis.


## LGBTQ 

$H_0$:Covid921 vaccine is not effective in LGBTQ.

$H_A$:Covid921 vaccine is effective in LGBTQ.

### Methods 

Treatment is CAT 
Sex is CAT 
So i use the CAT~CAT helper file



### Descriptive Results

#### Graphical Results

```{r}
barchartGC(~LGBTQ + treatment,data=FinalData)
barchartGC(~LGBTQ + treatment,data=FinalData, type="percent")

```

This shows that about an equal amount of the LBGTQ got covid rather or not the received the vaccine or placebo.

#### Numerical Results 

```{r}
table1 <- xtabs(~LGBTQ + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

This shows that about 50% of LGBTQ who recieved the placebo contracted covud 19, while about 50% of LBGTQ who recieved the covid 29 vaccine also contracted covid 19.


#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

P-value for chi-squared test is 0.1099 and fishers test is 0.1085 which is both smaller than 0.5 therefore we reject the null hypothesis.

## Druggies 


$H_A$:Covid921 v$H_0$:Covid921 vaccine is not effective in druggies.
accine is effective in druggies.

### Methods
 
The result is a categorical variable ~categorical variable 
Therefore i use CAT~CAT helper file.
 
 
### Descriptive Results

#### Graphical Results

```{r}
barchartGC(~DrugUser + treatment,data=FinalData)
barchartGC(~DrugUser + treatment,data=FinalData, type="percent")

```

This shows that about an equal amount of drugusers contracted covid 19 rather they received the vaccine or placebo.


#### Numerical Results 

```{r}
table1 <- xtabs(~DrugUser + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

This shows that aboyt 50% of drugusers who received placebo contracted covid 19 while about 50% of drugusers who got the vaccine contracted covid 19.

#### Inferential Results



```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for chi-square = , while the p-value for the fisher test= 0.8355 which is both smaller than 0.5 therefore we reject the null hypothesis.





# Overall Results and Conclusions 


$H_0$:Covid921 vaccine is not effective in males.

$H_A$:Covid921 vaccine is effective in males.

Treatment is CAT
Sex is CAT
So i use the CAT~CAT helper file

This table shows that males who got the vaccine are less likely to get covid than people who got placebo. 

This shows that 74% of males who took the placebo got covid 19 and 26% who took the vaccine got covid 19 as well. 


Both the chi and fisher test have the same p-value of p-value= 0.000000000000002 which is lower than 0.5 therefore we reject the null hypothesis.

$H_0$:Covid921 vaccine is not effective in females.

$H_A$:Covid921 vaccine is effective in females.

Treatment is CAT 
Sex is CAT 
So i use the CAT~CAT helper file

This shows that females who took the vaccine had a lower rate of getting covid 19 as opposed to those who took the placebo

his shows that 65% of females who took the placebo got covid 19 while 36% of females who got the vaccine comtracted covid 19.

The results of the p-value for chi and fishers test is p-value= 0.000000001247 which is lower than 0.5 therefore we reject the null hypothesis. 

$H_0$:Covid921 vaccine is not effective in LGBTQ.

$H_A$:Covid921 vaccine is effective in LGBTQ. 

The result is a categorical variable ~ categorical variable 
Therefore i use CAT~CAT helper file. 

This shows that about an equal amount of the LBGTQ got covid rather or not they received the vaccine or placebo.

This shows that about 50% of LGBTQ who recieved the placebo contracted covud 19, while about 50% of LBGTQ who recieved the covid 29 vaccine also contracted covid 19. 

P-value for chi-squared test is 0.1099 and fishers test is 0.1085 which is both smaller than 0.5 therefore we reject the null hypothesis. 

$H_A$:Covid921 v$H_0$:Covid921 vaccine is not effective in druggies.
accine is effective in druggies. 

The result is a categorical variable ~categorical variable 
Therefore i use CAT~CAT helper file.
 
This shows that about an equal amount of drugusers contracted covid 19 rather they received the vaccine or placebo. 

This shows that aboyt 50% of drugusers who received placebo contracted covid 19 while about 50% of drugusers who got the vaccine contracted covid 19. 

The p-value for chi-square = , while the p-value for the fisher test= 0.8355 which is both smaller than 0.5 therefore we reject the null hypothesis   




 

